---
title: "LDP Project Manuscript: How does browse severity influence understory plants?"
author:
- Jocelyn Biro^[Wilfrid Laurier University]
- Rod Kimble^[Hot Rod Stunts for Frank]
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
keywords: 
  - browse
  - moose
journal: "Moos-ecology"
date: "`r Sys.Date()`"
classoption: preprint, 2p, authoryear
bibliography: Biro_LDP_Project_References.bib
linenumbers: true
numbersections: true
csl: council-of-science-editors-author-date.csl
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)

```

# Introduction

Different levels of moose browsing pressure can change plant community compositions [@devriendtDelayedSuccessionAlternative2021].

# Methods

I used data from dryad (<https://datadryad.org/stash/dataset/doi:10.5061%2Fdryad.3xsj3txdp>). I calculated browse pressure (severity) by dividing the number of browsed twigs by the total number of twigs (proportion browsed). I used a scatterplot to investigate the total herbaceous plant cover in each plot (response variable) by the browse severity value of that plot (predictor variable). I repeated this for the sapling data (in number of saplings per site).

# Results

```{r, include=FALSE}
library(ggplot2)

#Read in cleaned csv file
understory.browse <- read.csv(here::here('data/clean/understory_browse.csv'))

#Create scatterplots
#Number of saplings x proportion of browse
p.sap <- ggplot(understory.browse, aes(x=total.sapling, y=prop.browse)) + geom_point()+labs(x="Total Saplings (#)", y="Proportion of Browse")

#Herb percent cover x proportion of browse
p.herb <- ggplot(understory.browse, aes(x=total.herb, y=prop.browse)) + geom_point()+labs(x="Total Herb Cover (%)", y="Proportion of Browse")
```

```{r fig1, fig.width = 5, fig.height = 3, fig.align='center', out.width="75%", fig.cap = "\\label{fig1}Number of saplings at different browse pressures", echo = FALSE}
p.sap
```

```{r fig2, fig.width = 5, fig.height = 3, fig.align='center', out.width="75%", fig.cap = "\\label{fig2}Herb percent covers at different browse pressures", echo = FALSE}
p.herb
```

\newpage

# Discussion

Yes, browsing pressure has an effect on understory plants [@devriendtDelayedSuccessionAlternative2021; @lorentzenkolstadPervasiveMooseBrowsing2018].

# References
