---
title: "LDP Project Manuscript"
author: "Jocelyn Biro"
date: "2022-09-16"
title: How does browse severity influence understory plants?
author:
  - name: Jocelyn Biro
    email: biro6470@mylaurier.ca
    affiliation: Wilfrid Laurier University
    correspondingauthor: true
    footnote: 1
  - name: Rod Kimble
    email: hotrod@fifteenbuses.com
    affiliation: Centre of Stunting for Frank's Surgery
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
abstract: |
  We looked at the understory plant cover under different levels of browse.
keywords: 
  - browse
  - moose
journal: "Moos-ecology"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: Biro_LDP_Project_References.bib
linenumbers: true
numbersections: true
csl: council-of-science-editors-author-date.csl
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
header-includes:
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# **How does browse severity influence understory plants?**

Jocelyn Biro^1^ , Rod Kimble^2^\*

1.  Department of Biology
2.  Centre of Stunting for Frank's Surgery

\newpage

# Abstract

We looked at understory plant covers under different levels of browse.

# Introduction

Different levels of moose browsing pressure can change plant community compositions [@devriendtDelayedSuccessionAlternative2021].

# Methods

I calculated browse pressure (severity) by dividing the number of browsed twigs by the total number of twigs (proportion browsed). I used a scatterplot to investigate the total herbaceous plant cover in each plot (response variable) by the browse severity value of that plot (predictor variable). I repeated this for the sapling data (in number of saplings per site).

# Results

```{r, include=FALSE}
library(ggplot2)

#Read in cleaned csv file
understory.browse <- read.csv('C:/Users/JocBi/OneDrive - Wilfrid Laurier University/Documents/CIEE-REPRO-PRO-2022/Biro_LDP_Project_2022/data/clean/understory_browse.csv')

#Create scatterplots
#Number of saplings x proportion of browse
p.sap <- ggplot(understory.browse, aes(x=total.sapling, y=prop.browse)) + geom_point()+labs(x="Total Saplings (#)", y="Proportion of Browse")

#Herb percent cover x proportion of browse
p.herb <- ggplot(understory.browse, aes(x=total.herb, y=prop.browse)) + geom_point()+labs(x="Total Herb Cover (%)", y="Proportion of Browse")
```

```{r fig1, fig.width = 5, fig.height = 3, fig.align='center', out.width="75%", fig.cap = "\\label{fig1}Number of saplings at different browse pressures", echo = FALSE}
p.sap
```

```{r fig1, fig.width = 5, fig.height = 3, fig.align='center', out.width="75%", fig.cap = "\\label{fig1}Herb percent covers at different browse pressures", echo = FALSE}
p.herb
```

# Discussion

Yes, browsing pressure has an effect on understory plants [@devriendtDelayedSuccessionAlternative2021, @lorentzenkolstadPervasiveMooseBrowsing2018].

# References
